\chapter*{Acknowledgments}

I almost quit three times. Here is why I did not.

I start with my parents, who raised us with curiosity. You brought our family across the world, let us run unsupervised through Venice, and taught us that adventure justifies risk. My father builds with his hands, always working on something in the garage. I build with code. I am the digital version of him. My mother was the voice of reason that I did not appreciate until later, the balance that kept our adventures from becoming disasters. Your investment in my education gave me the foundation upon which everything else was built, but it was that early freedom to explore that led me to Spain, chasing this doctorate.

To Dienke, my partner through the most challenging years of this journey. Our careers have been a game of ping pong: I went technical, you followed into data engineering, I chased remote work and data into AI, you scaled up, now you are circling back to AI. Few partners truly understand each other's work, but we speak the same language. For years, I talked about getting a job in the United States. You beat me to it. This dissertation would not exist without you pushing me forward, even as you race ahead.

My brother Pieter and I have followed suspiciously parallel paths: same school, same fraternity, same degree in mechanical engineering, same master's in artificial intelligence, and now both pursuing doctorates. He is essentially the improved model. He started his PhD in AI for medical imaging shortly after I began mine. Whether we are windsurfing, skiing, or writing papers, that sibling rivalry has been one of my greatest motivators. Thank you for making me better.

Among my friends, I have to single out Tumsa. Years ago, you built a bot that traded DogeCoin whenever Elon Musk tweeted about it. I joined that project. It led me to AI trading systems, then a master's thesis, then a publication at Columbia University, then this PhD at UPC, and finally this dissertation. The butterfly effect is real: a meme cryptocurrency bot led to a doctorate.

To my supervisors, Dr.~Angelos Antonopoulos and Dr.~Anna Umbert, who transformed me from a student into a researcher. Angelos, your vision exceeded my own in ways I did not always recognize. When I disappeared into details, you pulled me back to the larger picture. You gave me freedom to pursue ambitious ideas while keeping perspective on how they fit together, rescuing me from countless rabbit holes. Anna, your insights sharpened all aspects of this research. Together, you taught me that research is about asking the right questions, not just finding answers.

To Godfrey Kibalya, my colleague and co-author on every paper in this dissertation. Your thorough reviews caught what I missed, and our technical debates, sometimes heated but always productive, pushed this work further than I could have taken it alone.
