\section{System Model}
\label{sec:ch4_system_model}

Building on the generalization foundations established in Section~\ref{sec:ch2_generalization}, this section introduces the system model for our edge-cloud network, focusing on container resource traces' representation, heterogeneity, and workload forecasting using historical data.

\subsection{Traces}

We consider an edge-cloud network with containers operating at all infrastructure layers. Each container generates resource traces, represented as $T(t) \in \mathbb{R}^{d_t \times d_f}$, where $d_t$ denotes the time dimension and $d_f$ represents the number of features. These features include metrics such as \ac{CPU} usage, memory consumption, network bandwidth, and storage I/O operations. The matrix representation of the historical time-series data for a single trace is
\begin{equation}
T_i(t) = \begin{bmatrix}
d_{1,1} & d_{1,2} & \cdots & d_{1,d_t} \\
d_{2,1} & d_{2,2} & \cdots & d_{2,d_t} \\
\vdots & \vdots & \ddots & \vdots \\
d_{d_f,1} & d_{d_f,2} & \cdots & d_{d_f,d_t} 
\end{bmatrix} \in \mathbb{R}^{d_f \times d_t}.
\end{equation}

Let \( N_{\text{all}} \) denote the total number of container traces in the network. Hence, a set of container traces can be defined as
\begin{equation}
\mathbf{T}(t) = \begin{Bmatrix} T_1, T_2, \cdots, T_{N_\text{all}} \end{Bmatrix} \in \mathbb{R}^{d_f \times d_t \times N_{\text{all}}}.
\end{equation}

\subsection{Workload Forecasting}

Workload forecasting in edge-cloud networks predicts future resource usage based on historical container traces. The primary objective is to accurately forecast future resource usage, denoted as $\hat{T}$, for a given prediction length $d_y$, using historical data.

The forecasting model $f_\theta$, parameterized by $\theta$, is designed to make these predictions. Specifically, the model uses past observations of the resource usage trace $T_i$ to predict the future usage, given by
\begin{equation}
    \hat{T}_i[t + 1:t + d_y] = f_\theta\left(T_i[t - d_x + 1:t]\right),
\end{equation}
where $d_x$ represents the sequence length, indicating the number of past observations used for predictions, while $d_y$ represents the prediction length. The model parameters, denoted by $\theta$, are the variables that need to be optimized.

Our goal is to minimize the \ac{MAE} for each container trace in the network, i.e.,
\begin{align}
    &\min \frac{1}{d_y} \sum_{i=1}^{N_{\text{all}}} \left\| T_i[t + 1:t + d_y] - \hat{T}_i[t + 1:t + d_y] \right\|_1,
\end{align}
where $\hat{T}_i[t + 1:t + d_y]$ is the prediction vector and $T_i[t + 1:t + d_y]$ is the actual resource usage vector at time $t + d_y$ for trace $i$.

\subsection{Heterogeneity of Traces}
\label{sec:ch4_container_trace_heterogeneity}

The temporal variability and diversity of container traces require examination of resource usage pattern variations both within and across containers, arising from several factors:

\begin{itemize}
    \item \textbf{Workload Diversity:} Containers handle a wide range of workloads, each with different resource needs. To manage this diversity, we transform each time-series $T_i(t)$ into a feature space $\mathcal{Z}$, which allows us to group the workloads into $K$ distinct categories. This grouping helps us distinguish between different types of workloads both in space and time, which is essential for the solution we propose in Section~\ref{sec:ch4_proposed_solution}.
    
    \item \textbf{Temporal Variability:} Resource usage can show significant temporal variability due to factors such as fluctuating user demands, periodic tasks, and time-of-day effects. This variability can be modeled as $T_i(t) = \hat{T_i}(t) + \epsilon$, where $T_i(t)$ denotes the resource usage at time $t$ for trace $T_i$, $\hat{T_i}$ is the function capturing the predictable component, and $\epsilon$ represents stochastic variations.
\end{itemize}

To manage this heterogeneity, the model should be able to predict resource usage for unseen traces by considering trace diversity and temporal variability.
