% 4_research_contributions.tex
\section{Research Contributions}
\label{sec:ch1_research_contributions}

This section presents the peer-reviewed publications resulting from this dissertation research, organized by publication type. The chapter-publication mapping shows how each technical chapter draws from these published works.

\subsection*{Journal Publications}

\begin{itemize}

\item[{[J1]}] B. J. D. Gort, G. M. Kibalya and A. Antonopoulos, ``Distributed Agentic AI Framework for Autonomous Edge-Cloud Service Orchestration,'' \textit{\ac{IEEE} Transactions on Machine Learning in Communications and Networking}, Under review.

\item[{[J2]}] B. J. D. Gort, G. M. Kibalya and A. Antonopoulos, ``\ac{AgentEdge}: Agentic AI for Service Orchestration in the Edge-Cloud Continuum,'' \textit{\ac{IEEE} Communications Magazine}, Under review.

\item[{[J3]}] B. J. D. Gort, G. M. Kibalya and A. Antonopoulos, ``Attention-Driven AI Model Generalization for Workload Forecasting in the Compute Continuum,'' \textit{\ac{IEEE} Transactions on Machine Learning in Communications and Networking}, vol. 3, pp. 779-797, 2025, doi: 10.1109/TMLCN.2025.3584009.

\item[{[J4]}] B. J. D. Gort, G. M. Kibalya and A. Antonopoulos, ``\ac{AERO}: Adaptive Edge-Cloud Orchestration With a Sub-1K-Parameter Forecasting Model,'' \textit{\ac{IEEE} Transactions on Machine Learning in Communications and Networking}, vol. 3, pp. 463-478, 2025, doi: 10.1109/TMLCN.2025.3553100.

\item[{[J5]}] B. J. D. Gort, G. M. Kibalya, M. A. Serrano and A. Antonopoulos, ``Forecasting Trends in Cloud-Edge Computing: Unleashing the Power of Attention Mechanisms,'' \textit{\ac{IEEE} Communications Magazine}, vol. 63, no. 1, pp. 108-114, January 2025, doi: 10.1109/MCOM.001.2300583.
\end{itemize}

\subsection*{Conference Publications}

\begin{itemize}\sloppy
\item[{[C1]}] B. J. D. Gort, G. M. Kibalya, A. Umbert and A. Antonopoulos, ``Agentic AI with Emulation-assisted Prevalidation for Edge-Cloud Service Orchestration,'' \textit{2026 22nd \ac{IEEE} International Conference on Sensing, Communication, and Networking (SECON)}, Abu Dhabi, UAE, 2026, Under review.

\item[{[C2]}] B. J. D. Gort, G. M. Kibalya, A. Umbert and A. Antonopoulos, ``\ac{OmniFORE}: Attention-Based Generalization Framework for Edge-Cloud Workload Predictions,'' \textit{2025 \ac{IEEE} Symposium on Computers and Communications (ISCC)}, Bologna, Italy, 2025, pp. 1-6, doi: 10.1109/ISCC65549.2025.11326455.

\item[{[C3]}] B. J. D. Gort, G. M. Kibalya and A. Antonopoulos, ``Optimization of Distinct Time-Series Neural Architectures for Cloud-Edge Workload Prediction,'' \textit{2025 \ac{IEEE} International Conference on Machine Learning for Communication and Networking (ICMLCN)}, Barcelona, Spain, 2025, pp. 1-6, doi: 10.1109/ICMLCN64995.2025.11140432.
\end{itemize}

\subsection*{Book Chapter Publications}

\begin{itemize}
\item[{[B1]}] J. L. Berral-García, M. A. Serrano, P. Liu, J. Oliveras Torra, G. Bravo Rocca, M. Palacín Marfil, D. Buchaca, F. Agullo Lopez, P. Garcia, J. Guitart, R. Nou, B. J. D. Gort, L. Liatsas, G. M. Kibalya, A. Antonopoulos, D. Carrera and F. Guim, ``Intelligent orchestration mechanisms for image workload placement,'' in \textit{Advances in Image Processing, Reliability, and Artificial Intelligence}, M. J. Divan, P. Johri, F. Guim, D. Shchemelinin and M. Carranza, Eds. Elsevier, 2026, ch. 5, pp. 59-91, doi: 10.1016/B978-0-443-34266-0.00007-3.
\end{itemize}

\subsection*{Chapter-Publication Mapping}

The technical contributions are organized as follows, with each chapter grounded in peer-reviewed publications:

\begin{itemize}
    \item \textbf{Chapter~\ref{ch:adaptive-orchestration} (\ac{AERO}: Edge-Deployable Forecasting)}: Based primarily on [J4], with supporting work in [C3] on architecture optimization and validation methodology.
    
    \item \textbf{Chapter~\ref{ch:attention-driven-prediction} (\ac{OmniFORE}: Cross-Application Generalization)}: Based primarily on [J3], with supporting work in [J5] on attention mechanisms and [C2] on the generalization framework. Additional context on forecasting with transformers appears in [B1].
    
    \item \textbf{Chapter~\ref{ch:distributed-agentic-ai} (\ac{AgentEdge}: Autonomous Orchestration)}: Based primarily on [J1], with supporting work in [J2] and [C1] on emulation-based validation.
\end{itemize}

